Notes from the President

### 2013-12-28 pnsac/newsletter/article_EditorsNotes.tex
Source files for Dec13 newsletter.

This issue of the Chronicle carries the last of four chapters in
Richard Lodge's story of his time at Rolls-Royce. Taken together, they
make an interesting ethnology of an extinct industrial culture. The
whole series is accessible through our website,
{\color{blue}\texttt{\url{http://www.projectnorthstar.ca}}} (note that
the location of the neswletter archive will change when the new
website goes live) starting with the April 2013 issue of the
Chronicle.

@@

>>>>

This issue of the Chronicle carries the last of four chapters in
Richard Lodge's story of his time at Rolls-Royce. Taken together, they
make an interesting ethnology of an extinct industrial culture. The
whole series is accessible through our website,
{\color{blue}\texttt{\url{http://www.projectnorthstar.ca}}} (note that
the location of the newsletter archive will change when the new
website goes live) starting with the April 2013 issue of the
Chronicle.

\vspace{12pt}

@@

\vspace{12pt}

### 2013-12-28 pnsac/newsletter/article_Feature1.tex
Source files for Dec13 newsletter.

It is only now, actually working on the engine and referring
frequently to the exploded diagrams that I realize how complicated the
work of the Technical Publications Department must have been. The men
and women who worked there were largely unsung heroes in that it was
very tedious and unglamorous work which was of great importance for
the safe and trouble-free running of engines in service. I can
remember that I would hear from time to time of the difficulties of
ensuring that publication updates were actually delivered to many
remote places in the world and the considerable logistical
complications of ensuring that the packages arrived safely and
furthermore that the updates were actually inserted in the relevant
manuals, particularly as many of the engine users did not speak
English. I cannot remember now if any of the documentation was
translated into other languages although I have a recollection that we
employed people who could translate into French and Spanish.

@@

We were often dealing with the directors of R-R who lived on the top
floor of the Nightingale Road offices. Many of them were engineers and
they had little time for young bean counters. It was still a time of
optimism for the company. The huge contract for the Lockheed engines
had been signed and the company was still enjoying its reputation as
the company whose Merlin engines had saved Britain in the air
war. This was in 1965/6 and it was several years before our
projections of doom resulted in the bankruptcy of the
company. Fortunately I left the company in 1966 and was not involved
in what must have been a horrible process as a symbol of British
engineering excellence slowly went to the wall.

After the chaos caused by the demise of Rolls-Royce Limited, the
British government formed a new nationalized company to take over the
remains of the old one. This has been very much a success story and
the newly privatized Rolls-Royce is now a highly profitable and
innovative company. I look back on my days at R-R with nostalgia but I
now realize that much of the time we were all living in a fool's
paradise.

>>>>

It is only now, actually working on the engine and referring
frequently to the exploded diagrams that I realize how complicated the
work of the Technical Publications Department must have been. The men
and women who worked there were largely unsung heroes in that it was
very tedious and unglamorous work which was of great importance for
the safe and trouble free running of engines in service. I can
remember that I would hear from time to time of the difficulties of
ensuring that publication updates were actually delivered to many
remote places in the world and the considerable logistical
complications of ensuring that the packages arrived safely and
furthermore that the updates were actually inserted in the relevant
manuals, particularly as many of the engine users did not speak
English. I cannot remember now if any of the documentation was
translated into other languages although I have a recollection that we
employed people who could translate into French and Spanish.

@@

We were often dealing with the directors of R-R who lived on the top
floor of the Nightingale Road offices. Many of them were engineers and
they had little time for young bean counters. It was still a time of
optimism for the company. The huge contract for the Lockheed engines
had been signed and the company was still enjoying its reputation as
the company whose Merlin engines had helped to save Britain in the
Second World War. This was in 1965/6 and it was several years before
our projections of doom resulted in the bankruptcy of the
company. Fortunately I left the company in 1966 and was not involved
in what must have been a horrible process as a symbol of British
engineering excellence slowly went to the wall.

After the chaos caused by the demise of Rolls-Royce Limited, the
British government formed a new nationalized company to take over the
remains of the old one. This has been very much a success story and
the privatized Rolls-Royce is now a highly profitable and innovative
company. I look back on my days at R-R with nostalgia but I now
realize that much of the time we were all living in a fool's paradise.

### 2013-12-28 pnsac/newsletter/article_Feature2.tex
Source files for Dec13 newsletter.

\begin{figure}[htbp]
   \vspace{2em}
   \centering
   %name of the graphic, without the path AND in EPS format:
   \includegraphics[scale=0.5]{rcaf_c45.eps}
   %caption of the figure 
   \caption*{\small \em RCAF C45 Expeditor a military variant of the Beech 18.Purchased in large numbers during WW II and after for pilot training, navigator training, and as a utility transport.}
   %label of the figure, which has to correspond to \ref{}:
   \label{fig:rcaf_c45}
\end{figure}

@@

@@

>>>>

@@

% \begin{figure}[htbp]
%    \vspace{2em}
%    \centering
%    %name of the graphic, without the path AND in EPS format:
%    \includegraphics[scale=0.5]{rcaf_c45.eps}
%    %caption of the figure 
%    \caption*{\small \em RCAF C45 Expeditor a military variant of the Beech 18.Purchased in large numbers during WW II and after for pilot training, navigator training, and as a utility transport.}
%    %label of the figure, which has to correspond to \ref{}:
%    \label{fig:rcaf_c45}
% \end{figure}

@@

\end{multicols}

\begin{figure*}[htbp]
   \vspace{2em}
   \centering
   %name of the graphic, without the path AND in EPS format:
   \includegraphics[scale=1.0]{BEECHFLIGHT0001.eps}
   %caption of the figure 
   \caption*{\small \em Photo, taken by Tim, of two RCAF C45
     Expeditors (military variant of the Beech 18) during the ferry
     flight he describes in this article. These aircraft were
     purchased in large numbers during and after WW II for pilot
     training, navigator training, and as utility transports.}
   %label of the figure, which has to correspond to \ref{}:
   \label{fig:BEECHFLIGHT0001}
\end{figure*}

\begin{multicols}{2}

### 2013-12-28 pnsac/newsletter/article_PresReport.tex
Source files for Dec13 newsletter.

Canada Day is now behind us.  We had an excellent day with a good
number of our volunteers showing off the North Star to the general
public. Judging by the activity around the plane all the day, we have
now developed a display format which is interesting to visitors.
Canada Day was the only day this year when the plane was on display to
the general public.

On June 27, in co-operation with the Museum, who hosted the event, we
showed off the plane to a group of donors attending the Donor
Appreciation Evening.  This was a successful evening which is likely
to be repeated.

In my last Notes from the President, I referred to plans for
celebrating the 10th anniversary of the founding of the Project North
Star Association.  This should have taken place around June 24.
Unfortunately, the event passed without us being able mark the
day. This was mainly due to situations beyond our control.

Our 2013 AGM was held on June 8, and all the existing directors were
re-elected.  This will enable the directors, who have now become
experienced, to concentrate on long-term matters.

The actual work on the restoration of the plane is going very well, as
reported elsewhere in this newsletter. The major issue facing the
Board in the next year will be to consider how to secure the future of
PNSAC, both in terms of volunteers and the funding of restoration
work.

We have now reached a point where it is necessary to update our
Memorandum of Understanding (MOU) with the Museum and our work plan
for completing the restoration of the plane.  Our volunteers have
already worked over 60,000 hours on the project and we expect to work
a similar number of hours before the restoration is completed.

There have been several informal meetings to discuss our future
operations and the opportunities available for new volunteers.  Within
the next few weeks we are expecting to meet with the Director General
of the Aviation Museum and members of his staff to work on the MOU
update.

Later this month we are meeting with senior management of the Museum
Foundation, the body charged with corporate fundraising for the Canada
Science and Technology Museums Corporation, to discuss how PNSAC can
work with the Foundation to raise funds for the North Star restoration
and the museums in general.

With the initiatives above, we are hoping in the next few months to
consolidate our position as an important organization working with the
Aviation Museum

>>>>

As we come to the end of 2013, I think of it as an interesting year
with some disappointments and important successes. I believe we end
the year in a very positive way and I look forward to 2014 with
enthusiasm and renewed optimism.

First let me get the disappointments out of the way. 2013 was our 10th
anniversary year. We had hoped to make the event an opportunity to
show off our achievements during the first 10 years and create
additional interest around the plane. Unfortunately, due to events
beyond our control, it was not possible to stage a significant event
marking the anniversary. We therefore did our usual opening of the
plane to the public on Canada Day and marked the anniversary with our
PNSAC photographer taking photographs of members as they visited the
restored cockpit.

The second disappointment is more serious. During 2013 we did not take
on any new volunteers to work on the plane. We had several good
applicants but were unable to fit any of them into the working
schedule of volunteers at the Canada Aviation \& Space Museum
(CASM). This has two downsides; the first being that we are now
becoming short of active volunteers and the second being that we are
not training any new and sometimes younger volunteers to work on
pre-jet age aircraft. It is important to pass on the skills of the
older volunteers to younger members of our Association.

Funding the restoration of the aircraft is always a challenge and is
becoming more so as the Federal government cuts back on the funds it
appropriates to the CASM and other similar institutions..

You may well ask why I feel optimistic about 2014 after referring to
disappointments and problems. In the last two months the Directors of
PNSAC met with the Director General of CASM, Stephen Quick, and
subsequently a small group of directors met again with him together
with CASM conservation staff. The outcome of these meetings has been
an excellent exchange of views, a very firm commitment by the CASM to
continue to actively support the North Star restoration and a better
understanding by both sides of the challenges we face and possible
ways forward.

As many of you already know, volunteer work on the North Star will be
suspended for approximately the first three months of 2014. This is to
enable CASM to undertake a major reorganization of the Museum exhibits
in preparation for the centenary of the start of the First World
War. During this time the CASM professional staff will be occupied
with the reorganization and will not be able to provide the necessary
supervision of restoration work. During this time, though, we plan to
update the work plan and overall budget for completing the North Star
restoration, work on fundraising initiatives and finally update our
Memorandum of Understanding with CASM.

In addition to our hands-on restoration work, the Association has a
strong social side. Bill Tate, our Vice President, and his Special
Events team have organized some memorable bus trips, the last of which
was to Montreal in October. Bill has now turned his considerable
organizational skills to further developing our quarterly Members'
Meetings. In early December we had a great morning at CASM when
Capt. Bob Pearson of the Gimli Glider fame made a presentation to 150
of our members and members of other aviation organizations we invited
to attend. Although our Vice President has retired as an airline
captain, he certainly has not retired from anything else. Watch your
emails, Facebook and Twitter for news of events Bill and his team are
planning for 2014.
